---
title: "G7 Multi-country vs Single-country Comparison"
output: html_document
---

## Purpose

This file combines the pooled G7 model and the country-specific G7 models into one reusable workflow.

It does two things:

1. prints a country-level comparison table using the five anchors as one composition;
2. shows, for each G7 country, five anchor plots with three series: Observed, Multi-country, and Single-country.

The code below is based on the final pooled G7 model `vg_pr5` and the final country-specific G7 models stored in `g7_models`. In the pooled model, `vg_pr5` adds a `country:policyRate` interaction on top of smooth terms for headline inflation, policy rate, real GDP, and USD exchange rate, and it substantially improves AIC/BIC relative to `vg_pr4`. In the country-specific models, the G7 sample is restricted to 2014-01-01 through 2024-01-01; most countries use four smooth terms with `ps.int = 5`, Italy uses `ps.int = 4`, Japan keeps `ps.int = 5` with stronger response adjustment and `Maxit.outer = 400`, and the United States removes `usdExchangeRate` and uses three smooth terms with `ps.int = 6`. These choices come directly from your two project PDFs.  

## Assumed objects already available

This file assumes the following objects already exist in your session because they are created in your main scripts:

- `df2_10y`
- `vg_pr5`
- `g7_models`

If they are not already in memory, run your original model-building code first.

```{r setup, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
```

## 1. Basic settings

```{r}
comp <- c("bank", "govern", "growth", "market", "other")

g7_countries <- c(
  "Canada", "France", "Germany", "Italy",
  "Japan", "United Kingdom", "United States"
)

multi_model <- vg_pr5
```

## 2. Helper functions

```{r}
mae_fun <- function(obs, fit) {
  mean(abs(obs - fit), na.rm = TRUE)
}

rmse_fun <- function(obs, fit) {
  sqrt(mean((obs - fit)^2, na.rm = TRUE))
}

extract_fit_matrix <- function(model, prefix, comp_names = comp) {
  fit_mu <- as.matrix(fitted(model))
  out <- as.data.frame(fit_mu[, comp_names, drop = FALSE])
  names(out) <- paste0(comp_names, prefix)
  out
}
```

## 3. Build one comparison data set

This section avoids repeating extraction code in later chunks.

```{r}
# Multi-country fitted values
multi_fitted_data <- df2_10y %>%
  arrange(country, date_m) %>%
  bind_cols(extract_fit_matrix(multi_model, "_multi"))

# Single-country fitted values
single_fitted_list <- vector("list", length(g7_countries))
names(single_fitted_list) <- g7_countries

for (cty in g7_countries) {
  idx <- which(df2_10y$country == cty)
  ord <- order(df2_10y$date_m[idx])

  dat_cty <- df2_10y[idx[ord], ]
  dat_cty <- bind_cols(dat_cty, extract_fit_matrix(g7_models[[cty]], "_single"))

  single_fitted_list[[cty]] <- dat_cty
}

single_fitted_data <- bind_rows(single_fitted_list)

# Merge observed, pooled fitted, and country-specific fitted values
compare_df <- multi_fitted_data %>%
  select(
    country, date_m,
    all_of(comp),
    ends_with("_multi")
  ) %>%
  left_join(
    single_fitted_data %>%
      select(country, date_m, ends_with("_single")),
    by = c("country", "date_m")
  ) %>%
  arrange(country, date_m)
```

## 4. Composition-level error table

Here the five anchors are treated as one composition. Each country therefore gets four numbers:

- `multi_MAE`
- `multi_RMSE`
- `single_MAE`
- `single_RMSE`

```{r}
error_composition_country <- bind_rows(lapply(g7_countries, function(cty) {

  tmp <- compare_df %>%
    filter(country == cty) %>%
    arrange(date_m)

  obs_mat <- as.matrix(tmp[, comp])
  fit_multi_mat <- as.matrix(tmp[, paste0(comp, "_multi")])
  fit_single_mat <- as.matrix(tmp[, paste0(comp, "_single")])

  obs_vec <- as.vector(obs_mat)
  fit_multi_vec <- as.vector(fit_multi_mat)
  fit_single_vec <- as.vector(fit_single_mat)

  data.frame(
    country = cty,
    multi_MAE = mae_fun(obs_vec, fit_multi_vec),
    multi_RMSE = rmse_fun(obs_vec, fit_multi_vec),
    single_MAE = mae_fun(obs_vec, fit_single_vec),
    single_RMSE = rmse_fun(obs_vec, fit_single_vec)
  )
})) %>%
  arrange(country) %>%
  mutate(
    multi_MAE = round(multi_MAE, 4),
    multi_RMSE = round(multi_RMSE, 4),
    single_MAE = round(single_MAE, 4),
    single_RMSE = round(single_RMSE, 4)
  )

print(error_composition_country)
```

## 5. Plot helper

This plotting function returns a `ggplot` object, so it can be shown with `print()` in a loop and in knitted Markdown output.

```{r}
plot_country_5anchors <- function(cty_name, data_plot = compare_df) {

  tmp <- data_plot %>%
    filter(country == cty_name) %>%
    arrange(date_m)

  obs_long <- tmp %>%
    select(country, date_m, all_of(comp)) %>%
    pivot_longer(
      cols = all_of(comp),
      names_to = "anchor",
      values_to = "value"
    ) %>%
    mutate(model = "Observed")

  multi_long <- tmp %>%
    select(country, date_m, all_of(paste0(comp, "_multi"))) %>%
    rename_with(~ sub("_multi$", "", .x)) %>%
    pivot_longer(
      cols = all_of(comp),
      names_to = "anchor",
      values_to = "value"
    ) %>%
    mutate(model = "Multi-country")

  single_long <- tmp %>%
    select(country, date_m, all_of(paste0(comp, "_single"))) %>%
    rename_with(~ sub("_single$", "", .x)) %>%
    pivot_longer(
      cols = all_of(comp),
      names_to = "anchor",
      values_to = "value"
    ) %>%
    mutate(model = "Single-country")

  plot_df <- bind_rows(obs_long, multi_long, single_long) %>%
    mutate(
      anchor = factor(anchor, levels = comp),
      model = factor(model, levels = c("Observed", "Multi-country", "Single-country"))
    )

  ggplot(plot_df, aes(x = date_m, y = value, linetype = model, linewidth = model)) +
    geom_line() +
    facet_wrap(~ anchor, ncol = 2, scales = "free_y") +
    scale_linetype_manual(
      values = c(
        "Observed" = "dotted",
        "Multi-country" = "solid",
        "Single-country" = "longdash"
      )
    ) +
    scale_linewidth_manual(
      values = c(
        "Observed" = 0.4,
        "Multi-country" = 0.9,
        "Single-country" = 0.9
      )
    ) +
    labs(
      title = cty_name,
      x = "Date",
      y = "Proportion"
    ) +
    theme_bw() +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold", size = 16),
      strip.text = element_text(face = "bold"),
      legend.title = element_blank(),
      legend.position = "top"
    )
}
```

## 6. Show one example

```{r fig.width=11, fig.height=8}
p_canada <- plot_country_5anchors("Canada")
print(p_canada)
```

## 7. Show all G7 countries directly in the document

No PDF device is opened here. Each country plot is displayed directly using `print()`.

```{r fig.width=11, fig.height=8, results='asis'}
for (cty in g7_countries) {
  print(plot_country_5anchors(cty))
}
```

## 8. Optional: print the model-specific tuning summary

This small table keeps the final single-country specifications in one place.

```{r}
model_spec_table <- data.frame(
  country = g7_countries,
  n_smooth_terms = c(4, 4, 4, 4, 4, 4, 3),
  ps_int_used = c(5, 5, 5, 4, 5, 5, 6),
  special_note = c(
    "standard single-country version",
    "standard single-country version",
    "standard single-country version",
    "ps.int reduced to 4 for stability",
    "stronger response adjustment; Maxit.outer = 400",
    "standard single-country version",
    "usdExchangeRate removed"
  )
)

print(model_spec_table)
```
